---
title: "Milestone3"   
author: "Simon Chiu, Sabrina Tse, Sylvia Lee,  Hayley Boyce"  
date: '2019-04-10'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
```

## Objective 

Our team created the following analysis to address the following research question:

**"Does a person's choice of study location (home/academic setting/other public spaces) affect the time they take to finish their MDS assignments (exclusion of optional questions)?"**

Labs assignments comprise a big part of the Master of Data Science program. One characteristic of this program is the diversity of students that are enrolled in it. As students in this program, some of us live close to school, other live farther away and we all have different preferences when it comes to where we study. Consequentially, the members of our team have various study habits. As part of this study, we investigated if the choice of study location has an effect on the time a student takes to complete the weekly labs. Do people who choose to work at home perhaps get more distracted? or do they have more time to commit to the labs since they have less commuting time? To address this question we surveyed students in the 2018-2019 MDS cohort to obtain data to conduct our analysis. 

In the broader context, the completion and the quality of the lab assignments is crucial to a student's success in the program. This study is the beginning of accessing and exploring how external variables affects a student's enjoyment during the program and the amount of knowledge that they take away from the program. Exploration of these variables is worthwhile in developing strategies to help student succeed and improve the quality of the MDS program.  


## Method  

### Survey

To reiterate what we discussed in  Milestone 2, We conducted an anonymous online survey via Quatrics, a survey platform hosted by UBC. 

Survey link: https://ubc.ca1.qualtrics.com/jfe/form/SV_agz0I9HiBEeZqBL

The survey was conducted online anonymously during the lab session on April 3rd, 2019.  We expected that most of the participants to be UBC MDS students in the 2018-2019 cohort as the distribution was via the Slack channel for DSCI 554 course. The survey content focused on gathering MDS students' study behaviour and location preference in block 5. We limited the study scope to a specific recently completed block in the attempt to reduce the confounding variable of block difficulty. In the pursuit of more accurate results, we identified and addressed plausible confounding variables including commute time, procrastination tendencies, household responsibilities, and the completion of optional questions as part of the survey.  

In compliance to the [UBC Office of Research Ethics document on Using Online Surveys](https://ethics.research.ubc.ca/sites/ore.ubc.ca/files/documents/Online_Survey-GN.pdf), we did not collect any data that could be considered as direct or quasi-identifying to ensure the participants' anonymity. Additionally, the data is not available to the public. 


### The Data 

We gathered data from 59 participants from the MDS 2018-2019 cohort. 

The data is being stored in a private repository only accessible by the core study team and authorized personels. It can be located [here](https://github.ubc.ca/MDS-2018-19/DSCI_554-lab-time-analysis-DATA).

###### Table 1. Description of study parameters  

|Parameter             | type    | variable type | description                                     |
|:----------------------:|:--------:|:------:|:------------------------------------------------:|
| `Location`           |  fctr   | Categorical variable | a person's usual study location |
| `OptionalQ`          |  dbl    | Continuous variable | how long a person spends doing bonus questions | 
| `ProcrastLV`         |  int    | Discrete variable | the level of procrastination of which a person identifies. This is an ordinal scale from 1-7, 1 being not a procrastinator and 7 being the highest form of procrastination. | 
| `Household_Hr`       |  dbl    |  Continuous variable | the amount of daily household responsibilities in hours a person assumes.   |
| `Commute_Hm_Sch_Min` | dbl     | Continuous variable | the amount of time it takes to commute from their home to school (one way) in minutes | 
| `Commute_Stu_Loc_Min`| dbl     | Continuous variable | the amount of time it takes to move to their usual study location in minutes. | 
| `Time_On_Lab_Hr`     | dbl     | Continuous variable | the amount of time taken to complete all four labs in hours. | 
| `Spare_Time_Min`     | dbl     | Continuous variable | the amount of spare time a person has left before the submission time. | 



```{r,  echo= FALSE, include=FALSE}
# reading the data into Python or R and creating a summary table

data <- read_csv("../raw_data.csv")
```

```{r, warning = FALSE, echo= FALSE, include=FALSE}
# Cleaning and wrangling the data
#filtering out span and those who did not want to participate
data <- data %>% 
  filter( Status == "IP Address", Q0 == "I agree to participate" ) %>% 
  mutate(Q2 = ifelse(Q2 == "No", 0,  Q2_1_TEXT),
         Q15 = ifelse(Q15 == "Academic environment (school libraries, lounges, classrooms)", "Academic",  Q15),
         Q15 = ifelse(Q15 == "Other settings (coffee shops, etc)", "Other",  Q15)) %>% 
      select(Q15, Q2, Q3, Q4,Q16, Q11, Q5, Q18)


# adjusting the category types 
data <- data %>% 
  mutate(Q15 = as.factor(Q15), 
         Q2 = as.numeric(Q2), 
         Q3 = as.integer(Q3),  
         Q4 = as.numeric(Q4),  
         Q16 = as.numeric(Q16),  
         Q11 = as.numeric(Q11), 
         Q5 = as.numeric(Q5),
         Q18= as.numeric(Q18) )

data <- data %>% rename(Location = Q15, 
                        OptionalQ = Q2, 
                        ProcrastLV = Q3,  
                        Household_Hr = Q4,  
                        Commute_Hm_Sch_Min = Q16,  
                        Commute_Stu_Loc_Min = Q11, 
                        Time_On_Lab_Hr = Q5, 
                        Spare_Time_Min= Q18 ) 

```

<br>

###### Table 2. Sample dataset retrieved from survey response

```{r echo = FALSE, include = TRUE}
kable(head(data))

```

## Findings

###### Table 3. Count of students that prefer each study location

```{r echo=FALSE}
kable(data %>% group_by(Location) %>% summarize(n = n()))
```

We found that 24 students usually study in an academic setting, while 34 students prefer to study at home (Table 3.). We had one students that selected "other" study settings, we omitted the "other" location category from the rest of the analysis as it is not representative. 

 
#### EDA Analysis 

```{r echo = FALSE}
data_filter <- data %>% filter(Location != 'Other') 

plot_Q1 <- ggplot(data_filter, aes(x = Time_On_Lab_Hr)) + 
  geom_histogram( bins = 8, colour='white', fill = "#51B1D9")  + 
  theme_bw() + 
  labs(x= "Weekly average time (hours)", y = "Frequency", title = "Distribution of Time spent on labs Among MDS Students") +
  facet_wrap(~Location)

plot_Q1

```

###### Figure 1: Histogram of weekly average time spent on all labs in Block 5. 

Althought it is not exact, the distribution on amount of time spent on labs are roughly normal for both groups (Figure 1.). If the number of observations increased, we anticipate more normally distributed values. Given the small dataset, it seems to be reasonable to assume normality for both groups given the above histogram.

```{r echo = FALSE}
data_filter %>% ggplot(aes(Location, Time_On_Lab_Hr))+
  geom_violin(alpha = 0.3, 
                fill = "#51B1D9", 
                color = "#A9A9A9",
                weight = 0.3,
                draw_quantiles = c(0.25, 0.5, 0.75))+
  geom_jitter(width = 0.3, alpha = 0.3)+
  labs(x = 'Usual Study Location',
       y = 'Average weekly time (hr)',
       title = 'Average weekly time spent on lab assignments in relation to study location')
```

###### Figure 2: Average weekly time spent on labs in relation to the student's preferred study locations.

There is a larger range of time spent on labs by students whom study at school (Figure 2). The data appear much more spares and seem to have higher values which could be an indication that either people are getting distracted at school or people generally put more effort into the assignments. As seen in Figure 2, the sample distribution of the 'academic' group seems to have a greater skew then the sample distribution of students who study at home. Without addressing confounders, the centrality of the two distributions is very similar, which seems to suggest location has minimal effects on lab completion time.  

### Confounding Variables

In our proposal, we identified three confounding variables: procrastination, household responsibilities and the commute time to the study location. We first conducted statistical tests to determine confounding effects of these variables. 

#### Comparison - Baseline set-up

For comparison, we set up a baseline model by fitting a simple linear model on the dependent variable (Y) - Time spent on labs per week - and the independent variable (X) - the usual study location. Three groups included in the model are: the group who study in the academic settings("academic group"), the group who study at home (the home group) and the person who studies neither at home or at school (the "other" group).  In addition, 95% confidence intervals (CIs) were calculated using the baseline model. From the result below, we can see that the CI of both the home and other groups pass zero, implying that there is no significant difference in the time spent on labs when comparing the home and other groups to the academic group.

```{r echo = FALSE, include = TRUE}
lm_base <- lm(Time_On_Lab_Hr ~ Location, data = data)
lmsum <- summary(lm_base)
lmsum
```

```{r echo = FALSE, include = TRUE}
confint(lm_base)
```

We then added each confounding variable to the baseline model and refitted the model individually to observe the changes in the coefficients and standard errors. Out of the three confounding variables, we noticed the most significant difference in the coefficient and standard error comes from `commute time to the study location` calculated by the percentage changes. 

```{r echo = FALSE, include = TRUE}
lmsum_com_stu_loc <- summary(lm(Time_On_Lab_Hr ~ Location + Commute_Stu_Loc_Min, data = data))
lmsum_com_stu_loc
```

```{r echo = FALSE, include = TRUE}
print(paste0("LocationHome's Estimate change (%): ", (lmsum$coefficients[2, 1] - lmsum_com_stu_loc$coefficients[2, 1])/lmsum$coefficients[2, 1]*100, "%"))
print(paste0("LocationHome's Std. Err. change (%): ", (lmsum$coefficients[2, 2] - lmsum_com_stu_loc$coefficients[2, 2])/lmsum$coefficients[2, 2]*100, "%"))
print(paste0("LocationOther settings's Estimate change (%): ", (lmsum$coefficients[3, 1] - lmsum_com_stu_loc$coefficients[3, 1])/lmsum$coefficients[3, 1]*100, "%"))
print(paste0("LocationOther settings's Std. Err. change (%): ", (lmsum$coefficients[3, 2] - lmsum_com_stu_loc$coefficients[3, 2])/lmsum$coefficients[3, 2]*100, "%"))
```

To make better and more general inferences, we will only focus on the home and academic group in the following section as the other group only represents one individual.

#### The causal influence of commute time to the study location

The reference point in our model (intercept) is the academic group. On average, this group spends 28.7 hours per week on labs; The group who study at home (the 'home' group) comparatively spend 4.4 hours less.

After we introduced the confounding variable - `commute time to the study location per day` into the picture, the model revealed that the academic group spends 31.6 hours and the home group spends 5.6 hours less compared to the academic group.

From the observed changes, we can infer that the confounding variable indeed influences the relationship between time spent on labs per week and the study locations.  Our findings indicate that the longer time spent on travelling to a specific study location, the longer time the individual has to spend on completing the labs. 


## Conclusions 



## Assumptions

We are assuming that all participants responded truthfully and accurately. After performing EDA we are able to make the assumption that our variables are normally distributed and that residuals approximately follow a Normal distribution. We assume all our observations are normally distributed. 

## Limitations

The research we conducted was extremely limited. Our sample size was very small and therefore the ability to conduct causal inference was affected. We also had a single observation for the category "other" in the location parameter. This limited our analysis to addressing mostly the Academic and home locations. Other limitations include the categorization of the locations which included very few options. We could have extended to libraries, study room, and lounges or if people from home worked in an office or not. A major issue we encountered in that the interpretation of the question by the participants was varied. We believe that some of the outliers misread the questions. We had a limitation on candidates itself and potentially the responses could have been inaccurate or a poor representation of actual values since they were was an estimation according to the individuals. 

## Future Directions
  In the future, we might suggest conducting a similar analysis on larger samples, particularly undergraduate students, or specific larger classes. After further reflection, we believe age and collaboration with other students could have had confounding effects. Addition analysis could be done assessing if average study times were statistically significant in differing locations. Prediction based analysis instead of causal could also be an area of particular interest. 

## References 





