---
title: "Milestone3"   
author: "Simon Chiu, Sabrina Tse, Sylvia Lee,  Hayley Boyce"  
date: '2019-04-10'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
```

## Objective 

Our team created the following analysis to address the following research question:

**"Does a person's choice of study location (home/academic setting/other public spaces) affect the time they take to finish their MDS assignments (exclusion of optional questions)?"**

The members of our team have various study habits. Some of us live close to school, other live farther away and we all have different preferences when it comes to where we study. We wanted to know if the study location could have an effect on the time an individual takes to complete the weekly labs. Do people who choose to work at home perhaps get more distracted? or do they have more time to commit to the labs since they have no commuting time? To address this question we created a survey to obtain data to conduct our analysis. 


## Method  

### Survey

To reiterate what we discussed in  Milestone 2, We conducted an anonymous online survey. 
Survey link: https://ubc.ca1.qualtrics.com/jfe/form/SV_agz0I9HiBEeZqBL


The survey was conducted online anonymously during the lab session on April 3rd, 2019.  We expected that most of the participants to be UBC MDS students in the 2018-2019 cohort since the survey content focused on gathering MDS students' study behaviour and location preference.


### The Data 

Our survey had 59 participants from the MDS program answer questions regarding lab completion times specifically for Block 5. In the pursuit of more accurate results, we tried to reduce the confounding variable of block difficulty by limiting to a specific recently completed block. To keep complete anonymity, we did not collect any data that could be considered as direct or quasi-identifying. 

The data is being stored in a private repository only accessible by key personel and those with authorization. It can be located [here](https://github.ubc.ca/MDS-2018-19/DSCI_554-lab-time-analysis-DATA)

For clarity the parameters are defined below:  


|Parameter             | type    | variable type | description                                     |
|:----------------------:|:--------:|:------:|:------------------------------------------------:|
| `Location`           |  fctr   | Categorical variable | a person's usual study location |
| `OptionalQ`          |  dbl    | Continuous variable | how long a person spends doing bonus questions | 
| `ProcrastLV`         |  int    | Discrete variable | the level of procrastination of which a person identifies. This is an ordinal scale from 1-7, 1 being not a procrastinator and 7 being the highest form of procrastination. | 
| `Household_Hr`       |  dbl    |  Continuous variable | the amount of daily household responsibilities in hours a person assumes.   |
| `Commute_Hm_Sch_Min` | dbl     | Continuous variable | the amount of time it takes to commute from their home to school (one way) in minutes | 
| `Commute_Stu_Loc_Min`| dbl     | Continuous variable | the amount of time it takes to move to their usual study location in minutes. | 
| `Time_On_Lab_Hr`     | dbl     | Continuous variable | the amount of time taken to complete all four labs in hours. | 
| `Spare_Time_Min`     | dbl     | Continuous variable | the amount of spare time a person has left before the submission time. | 



```{r,  echo= FALSE, include=FALSE}
# reading the data into Python or R and creating a summary table

data <- read_csv("../raw_data.csv")
```

```{r, warning = FALSE, echo= FALSE, include=FALSE}
# Cleaning and wrangling the data
#filtering out span and those who did not want to participate
data <- data %>% 
  filter( Status == "IP Address", Q0 == "I agree to participate" ) %>% 
  mutate(Q2 = ifelse(Q2 == "No", 0,  Q2_1_TEXT),
         Q15 = ifelse(Q15 == "Academic environment (school libraries, lounges, classrooms)", "Academic",  Q15)) %>% 
      select(Q15, Q2, Q3, Q4,Q16, Q11, Q5, Q18)


# adjusting the category types 
data <- data %>% 
  mutate(Q15 = as.factor(Q15), 
         Q2 = as.numeric(Q2), 
         Q3 = as.integer(Q3),  
         Q4 = as.numeric(Q4),  
         Q16 = as.numeric(Q16),  
         Q11 = as.numeric(Q11), 
         Q5 = as.numeric(Q5),
         Q18= as.numeric(Q18) )

data <- data %>% rename(Location = Q15, 
                        OptionalQ = Q2, 
                        ProcrastLV = Q3,  
                        Household_Hr = Q4,  
                        Commute_Hm_Sch_Min = Q16,  
                        Commute_Stu_Loc_Min = Q11, 
                        Time_On_Lab_Hr = Q5, 
                        Spare_Time_Min= Q18 ) 

```


```{r echo = FALSE, include = TRUE}
kable(head(data))
  
```


## Findings
 
 
 

## Conclusions 



## Assumptions

We are assuming that all participants responded truthfully and accurately. After performing EDA we are able to make the assumption that our variables are normally distributed and that residuals approximately follow a Normal distribution. We assume all our observations are normally distributed. 

## Limitations

The research we conducted was extremely limited. Our sample size was very small and therefore the ability to conduct causal inference was affected. We also had a single observation for the category "other" in the location parameter. This limited our analysis to addressing mostly the Academic and home locations. Other limitations include the categorization of the locations which included very few options. We could have extended to libraries, study room, and lounges or if people from home worked in an office or not. A major issue we encountered in that the interpretation of the question by the participants was varied. We believe that some of the outliers misread the questions. We had a limitation on candidates itself and potentially the responses could have been inaccurate or a poor representation of actual values since they were was an estimation according to the individuals. 

## Future Directions
  In the future, we might suggest conducting a similar analysis on larger samples, particularly undergraduate students, or specific larger classes. After further reflection, we believe age and collaboration with other students could have had confounding effects. Addition analysis could be done assessing if average study times were statistically significant in differing locations. Prediction based analysis instead of causal could also be an area of particular interest. 

## References 





